---
title: "Getting Started with rmdreportdeck"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with rmdreportdeck}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Overview

`rmdreportdeck` provides:

- HTML rendering with a collapsible report shell
- PDF rendering with the same `.runinfo` provenance sidecar
- helper functions for embedding downloadable figure and data assets in HTML reports
- loop-aware item panels for repeated outputs created inside `for` loops
- thin POSIX CLI wrappers for `knit2html` and `knit2pdf`

# Authoring model

Write a standard `.Rmd` file and keep the structure simple:

- use the YAML header normally
- use `#` headings for major sections
- use `##` headings for subsections
- include a `# Goal`, `# Description`, `# Methods`, `# Workflow`, or `# Pipeline`
  section when you want that section to open by default
- keep the analytical content in ordinary chunks; `rmdreportdeck` styles the output shell rather than changing normal R Markdown semantics
- when a subsection contains many repeated outputs, keep one `##` heading and render the repeated items with `report_item_panel()` and `report_loop_section()`

For HTML reports, the shell behavior is:

- H1 sections become collapsible top-level panels
- H2 sections become collapsible nested panels
- the sidebar is generated from the section hierarchy
- key figures and tables can include in-page download links

# Minimal HTML pattern

````markdown
---
title: "P01 | Example Report"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(knitr)
library(rmdreportdeck)

df <- data.frame(group = c("A", "B"), value = c(2, 5))
plot_obj <- ggplot(df, aes(group, value, fill = group)) + geom_col()
```

# Goal

Describe the purpose of the report here.

## Figure

```{r}
print(report_asset_bar(
  "Downloads",
  report_download_link("PNG", "plot.png", report_plot_png_data_uri(plot_obj)),
  report_download_link("PDF", "plot.pdf", report_plot_pdf_data_uri(plot_obj)),
  report_download_link("TSV", "plot.tsv", report_text_data_uri(report_tsv_text(df)))
))
print(plot_obj)
```

## Table

```{r}
knitr::kable(df)
```
````

# Loop-generated outputs

For repeated outputs such as one figure per sample or pathway, use one H2 subsection and build loop items inside a regular R chunk.

````markdown
## Per-sample QC

```{r results='asis'}
items <- vector("list", nrow(samples))

for (i in seq_len(nrow(samples))) {
  row <- samples[i, , drop = FALSE]
  plot_obj <- ggplot(row, aes(sample, value, fill = sample)) + geom_col()

  items[[i]] <- report_item_panel(
    title = paste0("Sample ", row$sample),
    asset_bar = report_asset_bar(
      "Downloads",
      report_download_link("PNG", paste0("sample-", row$sample, ".png"), report_plot_png_data_uri(plot_obj)),
      report_download_link("PDF", paste0("sample-", row$sample, ".pdf"), report_plot_pdf_data_uri(plot_obj)),
      report_download_link("TSV", paste0("sample-", row$sample, ".tsv"), report_text_data_uri(report_tsv_text(row)))
    ),
    note = "Synthetic example item.",
    report_item_plot(plot_obj),
    report_item_table(row)
  )
}

report_loop_section(items)
```
````

# Rendering

From R:

```{r eval=FALSE}
render_html_report_with_runinfo("report.Rmd")
render_pdf_report_with_runinfo("report.Rmd")
```

From the CLI wrappers:

```bash
knit2html report.Rmd param=value
knit2pdf report.Rmd param=value
```

Both renderers write a `.runinfo` sidecar next to the rendered report.

# Examples

The package also ships generic example reports:

- `system.file("examples", "minimal_html_report.Rmd", package = "rmdreportdeck")`
- `system.file("examples", "loop_html_report.Rmd", package = "rmdreportdeck")`
- `system.file("examples", "minimal_pdf_report.Rmd", package = "rmdreportdeck")`
