---
title: "Getting Started with rmdreportdeck"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with rmdreportdeck}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Overview

`rmdreportdeck` provides:

- HTML rendering with a collapsible report shell
- PDF rendering with the same `.runinfo` provenance sidecar
- helper functions for embedding downloadable figure and data assets in HTML reports
- higher-level bundle helpers for the common “plot + downloadable source table” pattern
- standardized asset naming helpers for block-oriented reports
- loop-aware item panels for repeated outputs created inside `for` loops
- thin POSIX CLI wrappers for `knit2html` and `knit2pdf`

# Authoring model

Write a standard `.Rmd` file and keep the structure simple:

- use the YAML header normally
- use `#` headings for major sections
- use `##` headings for subsections
- include a `# Goal`, `# Description`, `# Methods`, `# Workflow`, or `# Pipeline`
  section when you want that section to open by default
- keep the analytical content in ordinary chunks; `rmdreportdeck` styles the output shell rather than changing normal R Markdown semantics
- when a subsection contains many repeated outputs, keep one `##` heading and render the repeated items with `report_item_panel()` and `report_loop_section()`

For HTML reports, the shell behavior is:

- H1 sections become collapsible top-level panels
- H2 sections become collapsible nested panels
- the sidebar is generated from the section hierarchy
- key figures and tables can include in-page download links

The package has two download-helper layers:

- low-level primitives such as `report_download_link()`, `report_plot_png_data_uri()`, and `report_text_data_uri()`
- a convenience bundle layer built around `report_plot_bundle()`, `report_bundle_download_links()`, `report_bundle_asset_bar()`, and `report_tsv_download_link()`

It also provides a naming layer for reports that use block identifiers such as
`V01.03`:

- `report_asset_name()`
- `report_figure_name()`
- `report_table_name()`

Use the bundle layer when one plot and one data frame belong together and you
do not need custom download assembly logic.

# Minimal HTML pattern

````markdown
---
title: "P01 | Example Report"
output: html_document
params:
  prjpath: !expr paste0(normalizePath(getwd()), "/")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)

library(ggplot2)
library(knitr)
library(rmdreportdeck)
reportdeck_setup()

df <- data.frame(group = c("A", "B"), value = c(2, 5))
plot_obj <- ggplot(df, aes(group, value, fill = group)) + geom_col()
bundle <- report_plot_bundle(plot_obj, block_prefix = "V01.01", suffix = "plot", data = df)
```

# Goal

Describe the purpose of the report here.

## Figure

```{r fig_assets, results='asis', echo=FALSE}
cat(as.character(report_bundle_asset_bar("Downloads", bundle, tsv_label = "TSV")))
```

```{r fig}
print(plot_obj)
```

## Table

```{r table}
knitr::kable(df)
```
````

Two conventions to follow in every HTML report:

- **`reportdeck_setup()`**: call it immediately after `library(rmdreportdeck)` in the
  setup chunk. It configures knitr caching with provenance tracking and
  registers size-aware hooks. Without it, caching is not coordinated with the
  report shell.

- **`results='asis'` + `cat(as.character(...))`**: asset bars and other HTML
  widgets must be emitted in chunks with `results='asis'`. Use
  `cat(as.character(...))` rather than `print()`. Keeping the asset bar in its
  own chunk (separate from the plot chunk) makes caching and chunk reuse
  cleaner.

# Loop-generated outputs

For repeated outputs such as one figure per sample or pathway, use one H2 subsection and build loop items inside a regular R chunk.

````markdown
## Per-sample QC

```{r per_sample_qc, results='asis', echo=FALSE}
items <- vector("list", nrow(samples))

for (i in seq_len(nrow(samples))) {
  row <- samples[i, , drop = FALSE]
  plot_obj <- ggplot(row, aes(sample, value, fill = sample)) + geom_col()
  bundle <- report_plot_bundle(
    plot_obj,
    block_prefix = paste0("V01.", sprintf("%02d", i)),
    suffix = row$sample,
    width = 5,
    height = 3.5,
    data = row
  )

  items[[i]] <- report_item_panel(
    title = paste0("Sample ", row$sample),
    asset_bar = report_bundle_asset_bar(
      "Downloads",
      bundle,
      tsv_label = "TSV"
    ),
    note = "Synthetic example item.",
    report_item_plot(plot_obj),
    report_item_table(row)
  )
}

cat(as.character(report_loop_section(items)))
```
````

The loop chunk must have `results='asis'` so the HTML panels are injected
directly into the document. `cat(as.character(...))` is the correct emission
idiom — `print()` will not work for HTML tag objects.

# Bundles and low-level helpers

The bundle helpers do not replace the existing low-level API; they package it.

Use the low-level helpers when:

- you already have custom link logic
- you need non-TSV text payloads
- a download bar mixes several unrelated files

Use the bundle helpers when:

- one plot and one data frame travel together
- the report should optionally write `.pdf`, `.svg`, or `.tsv` side files while still embedding in-page downloads
- the same per-item pattern repeats inside a loop

```{r eval=FALSE}
bundle <- report_plot_bundle(
  plot_obj,
  block_prefix = "V01.03",
  suffix = "figure",
  width = 8,
  height = 6,
  data = df,
  plot_dir = "results/plots",
  data_dir = "results/data",
  save_svg = TRUE
)

report_bundle_asset_bar("Figure 1", bundle, tsv_label = "Source TSV")
report_tsv_download_link(df, "table.tsv", "Table TSV")
```

# Standardized asset naming

Use the naming helpers when the report has a section or block prefix and the
same name should drive filenames, download controls, and visible labels.

The naming convention is:

`<asset_type><block_prefix>[.<suffix>][-<descriptor>]`

```{r eval=FALSE}
block_prefix <- "V01.03"

figure_stem <- report_figure_name(block_prefix, "nGenes_per_sample")
table_name <- report_table_name(block_prefix, "nGenes_per_sample.tsv")
aux_name <- report_asset_name("Aux", block_prefix, "qc", "raw")
```

Typical report usage:

````markdown
```{r}
block_prefix <- "V01.03"
figure_stem <- report_figure_name(block_prefix, "nGenes_per_sample")
figure_pdf <- paste0(figure_stem, ".pdf")
table_name <- report_table_name(block_prefix, "nGenes_per_sample.tsv")

ggsave(file.path("results/plots", figure_pdf), plot_obj, width = 8, height = 6)
write.table(df, file.path("results/data", table_name), sep = "\t", quote = FALSE, row.names = FALSE)

print(report_asset_bar(
  figure_stem,
  report_download_link("Download PNG", paste0(figure_stem, ".png"), png_uri),
  report_download_link("Download PDF", figure_pdf, pdf_uri)
))
```
````

The bundle helpers already use the naming policy internally. Use the naming
helpers directly when you are not creating a bundle.

# Rendering

From R:

```{r eval=FALSE}
render_html_report_with_runinfo("report.Rmd")
render_pdf_report_with_runinfo("report.Rmd")
```

From a project-level wrapper script:

```bash
scripts/knit2html report.Rmd param=value
```

Both renderers write a `.runinfo` sidecar next to the rendered report.

## Project-level `knit2html` wrapper

Projects using `renv` need a thin wrapper at `scripts/knit2html` in the
project root. The wrapper must:

1. Resolve the Rmd path to absolute before changing directory.
2. `cd` to the project root so `renv` activates via `.Rprofile`.
3. Set `R_PROFILE_USER` explicitly so the right profile is sourced even when
   R is invoked non-interactively.
4. Call `rmdreportdeck::knit2html_cli()` to do the actual rendering.

Minimal template:

```bash
#!/usr/bin/env bash
set -euo pipefail

if [ $# -lt 1 ]; then
  echo "Usage: knit2html <Rmdfile> [param=value ...]" > /dev/stderr; exit 1
elif [ ! -e "$1" ]; then
  echo "File does not exist: $1" > /dev/stderr; exit 1
fi

SCRIPT_DIR="$( cd "$( dirname "${BASH_SOURCE[0]}" )" && pwd )"
PROJECT_ROOT="$( cd "$SCRIPT_DIR/.." && pwd )"
RMD_FILE="$1"; shift

if [[ "$RMD_FILE" != /* ]]; then RMD_FILE="$PWD/$RMD_FILE"; fi

cd "$PROJECT_ROOT"
export R_PROFILE_USER="$PROJECT_ROOT/.Rprofile"

Rscript -e 'rmdreportdeck::knit2html_cli(commandArgs(trailingOnly = TRUE))' \
  "$RMD_FILE" "$@"
```

Projects may extend this to source environment config files or to support
loading `rmdreportdeck` from source via a `REPORTDECK_DEV_PATH` variable.

# Examples

The package also ships generic example reports:

- `system.file("examples", "minimal_html_report.Rmd", package = "rmdreportdeck")`
- `system.file("examples", "loop_html_report.Rmd", package = "rmdreportdeck")`
- `system.file("examples", "minimal_pdf_report.Rmd", package = "rmdreportdeck")`
